---
title: "Midterm 1 W24"
author: "Kassidy Yu"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

## Instructions

Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code must be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above.

Your code must knit in order to be considered. If you are stuck and cannot answer a question, then comment out your code and knit the document. You may use your notes, labs, and homework to help you complete this exam. Do not use any other resources- including AI assistance.

Don't forget to answer any questions that are asked in the prompt!

Be sure to push your completed midterm to your repository. This exam is worth 30 points.

## Background

In the data folder, you will find data related to a study on wolf mortality collected by the National Park Service. You should start by reading the `README_NPSwolfdata.pdf` file. This will provide an abstract of the study and an explanation of variables.

The data are from: Cassidy, Kira et al. (2022). Gray wolf packs and human-caused wolf mortality. [Dryad](https://doi.org/10.5061/dryad.mkkwh713f).

## Load the libraries

```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the wolves data

In these data, the authors used `NULL` to represent missing values. I am correcting this for you below and using `janitor` to clean the column names.

```{r message=FALSE, warning=FALSE}
wolves <- read.csv("data/NPS_wolfmortalitydata.csv", na = c("NULL")) %>% clean_names()
```

## Questions

Problem 1. (1 point) Let's start with some data exploration. What are the variable (column) names?\
The column names are "park", "biolyr", "pack", "packcode", "packsize_aug", "mort_yn", "mort_all", "mort_lead", "mort_nonlead", "reprody1", and "persisty1".

```{r}
colnames(wolves)
```

Problem 2. (1 point) Use the function of your choice to summarize the data and get an idea of its structure.

```{r}
glimpse(wolves)
```

Problem 3. (3 points) Which parks/ reserves are represented in the data? Don't just use the abstract, pull this information from the data.\
The parks represented in the data are "DENA", "GNTP", "VNP", "YNP", and "YUCH".

```{r}
park_factors <- as.factor(wolves$park)
levels(park_factors)
```

Problem 4. (4 points) Which park has the largest number of wolf packs?\
DENA has the largest number of wolf packs at 69 total.

```{r}
wolves %>%
  group_by(park) %>%
  summarize(num_packs = n_distinct(pack)) %>%
  filter(num_packs == max(num_packs))
```

Problem 5. (4 points) Which park has the highest total number of human-caused mortalities `mort_all`?\
The Yukon-Charley Rivers National Preserve ("YUCH") has the highest total number of human-caused mortalities.

```{r}
wolves %>%
  group_by(park) %>%
  summarize(human_mortality_total = sum(mort_all)) %>%
  filter(human_mortality_total == max(human_mortality_total))
```

The wolves in [Yellowstone National Park](https://www.nps.gov/yell/learn/nature/wolf-restoration.htm) are an incredible conservation success story. Let's focus our attention on this park.

Problem 6. (2 points) Create a new object "ynp" that only includes the data from Yellowstone National Park.

```{r}
ynp = wolves %>%
  filter(park == "YNP")
```

Problem 7. (3 points) Among the Yellowstone wolf packs, the [Druid Peak Pack](https://www.pbs.org/wnet/nature/in-the-valley-of-the-wolves-the-druid-wolf-pack-story/209/) is one of most famous. What was the average pack size of this pack for the years represented in the data?\
The average pack size of the Druid Peak Pack for the years represented is 13.93333.

```{r}
druid <- ynp %>%
  filter(pack == "druid") %>%
  arrange(biolyr)
mean(druid$packsize_aug)
```

Problem 8. (4 points) Pack dynamics can be hard to predict- even for strong packs like the Druid Peak pack. At which year did the Druid Peak pack have the largest pack size? What do you think happened in 2010?\
The Druid Peak pack had the largest pack size of 37 in 2001.

```{r}
druid %>%
  select(pack, biolyr, packsize_aug)  %>%
  filter(packsize_aug == max(packsize_aug))
```

It seems like the pack died out in 2010 - there was no human-caused mortality around 2010, so perhaps there was a famine.

```{r}
druid %>%
  select(pack, biolyr, packsize_aug, mort_all) %>%
  arrange(biolyr)
```

Problem 9. (5 points) Among the YNP wolf packs, which one has had the highest overall persistence `persisty1` for the years represented in the data? Look this pack up online and tell me what is unique about its behavior- specifically, what prey animals does this pack specialize on?\
The Mollie's pack had the highest overall persistence. They have had female alphas with long reigns, which has allowed for a stable pack life. Their unity allows them to excel in hunting down bison, which are the hardest prey for wolves to kill. They use snow, which weakens bison, to their advantage. Sources: [Greater Yellowstone Coalition](https://greateryellowstone.org/blog/2020/studyingwolves) and [The Spokesman-Review](https://www.spokesman.com/stories/2012/jan/15/hungry-wolf-pack-rearranges-balance-in/).

```{r}
ynp %>%
  group_by(pack) %>%
  filter(persisty1 == 1) %>%
  select(pack, persisty1) %>%
  count(pack) %>%
  arrange(desc(n))
```

Problem 10. (3 points) Perform one analysis or exploration of your choice on the `wolves` data. Your answer needs to include at least two lines of code and not be a summary function.\
I selected Yukon-Charley Rivers National Preserve, since in Problem 5, I found that they had the highest total human-caused mortality. Then, I wanted to see which year and packs these mortality numbers, so I selected the variables of interest and sorted by the highest `mort_all`. I found that the highest `mort_all` for one year in YUCH was in 2012, in which 24 wolves from the 70 Mile pack were killed by humans.

```{r}
wolves %>%
  filter(park == "YUCH", mort_all != 0) %>%
  select(biolyr, pack, mort_all) %>%
  arrange(desc(mort_all))
```
