---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?\
AI, or artificial intelligence, refers to the simulation of human intelligence processes by machines, especially computer systems. These processes include learning (the acquisition of information and rules for using the information), reasoning (using rules to reach approximate or definite conclusions), and self-correction. AI encompasses various subfields, such as machine learning, natural language processing, computer vision, robotics, and expert systems, among others. The ultimate goal of AI is to create systems that can perform tasks that would typically require human intelligence. These tasks may include problem-solving, decision-making, perception, understanding natural language, and even creativity.

2. What is ChatGPT?\
ChatGPT is a conversational AI developed by OpenAI, based on the GPT (Generative Pre-trained Transformer) architecture. It's designed to generate human-like text responses given a prompt or input from a user. ChatGPT is capable of understanding and generating text in natural language, allowing it to engage in conversations, answer questions, provide information, offer suggestions, and more.

The model is trained on a vast dataset comprising various sources of text from the internet, which enables it to have a broad understanding of language and a wide range of topics. ChatGPT utilizes deep learning techniques, particularly transformer models, to process and generate text based on the patterns and structures it has learned during training.

ChatGPT has been used in various applications, including customer service chatbots, language translation tools, writing assistance software, and virtual assistants. It's continuously being improved and updated by OpenAI to enhance its capabilities and performance.

3. What is Github Copilot?\
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It integrates directly into code editors, such as Visual Studio Code, and provides suggestions and completions for code as developers type. GitHub Copilot is built on OpenAI's Codex, which is a language model trained on a vast amount of code from open-source repositories, documentation, and other sources.

Using natural language prompts and context from the code being written, GitHub Copilot generates relevant code snippets, function definitions, and even entire functions. It aims to assist developers in writing code more efficiently by reducing the need to manually type repetitive or boilerplate code and by providing helpful suggestions based on the task at hand.

GitHub Copilot has garnered attention for its ability to quickly generate code that aligns with the developer's intentions and for its potential to speed up the development process. However, it's important for developers to review and understand the suggestions provided by GitHub Copilot to ensure the correctness and security of the code being written.

4. What is R?\
R is a programming language and software environment primarily used for statistical computing and graphics. It provides a wide variety of statistical and graphical techniques, including linear and nonlinear modeling, time-series analysis, classification, clustering, and more. R is highly extensible, with a vast ecosystem of packages contributed by the R community, which cover diverse domains such as machine learning, data visualization, bioinformatics, and econometrics.

R is both a programming language and an interactive environment, making it well-suited for data analysis and exploration. It features a rich set of tools for data manipulation, transformation, and visualization, allowing users to analyze and visualize data effectively.

One of the key strengths of R is its open-source nature, which has led to its widespread adoption in academia, research, and industry. It is commonly used by statisticians, data scientists, researchers, and analysts for tasks such as data cleaning, modeling, hypothesis testing, and generating reports.

R's syntax is designed to be expressive and concise, making it easy to work with data structures such as vectors, matrices, data frames, and lists. Additionally, R has a vibrant community and extensive documentation, making it accessible to users of varying skill levels.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("data/netflix_titles.csv")

```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*
I had to ask for R 

## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
# Check the structure of the data frame
str(netflix)

# Summarize the data frame
summary(netflix)

# Display the first few rows of the data frame
#head(netflix)

# Display the last few rows of the data frame
#tail(netflix)

# Get dimensions of the data frame
#dim(netflix)

# Get names of columns in the data frame
#colnames(netflix)

# Get the class of each column in the data frame
#sapply(netflix, class)

# Get unique values in a specific column
#unique(netflix$column_name)

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*
ChatGPT gave me lots of options

Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
# Filter movies from the 'netflix' data frame
netflix_movies <- netflix[netflix$type == "Movie", ]

# Display the first few rows of the 'netflix_movies' data frame
head(netflix_movies)
```


Which director has the most movies on netflix?
```{r}
director_counts <- netflix_movies %>%
  tabyl(director)
director_counts #there are some movies with multiple directors

# Sort the table in descending order of movie counts
#director_counts <- director_counts[order(-director_counts$n), ]

# Display the director with the most movies
#director_counts[1, ]
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Make sure to separate movies with multiple directors, separated by a comma
```{r}
# Separate movies with multiple directors into separate rows
netflix_movies_separated <- netflix_movies %>%
  separate_rows(director, sep = ", ")  # Split directors separated by comma

# Create a frequency table of directors and their movie counts
director_counts <- netflix_movies_separated %>%
  tabyl(director)

# Sort the table in descending order of movie counts
director_counts <- director_counts[order(-director_counts$n), ]

# Display the director with the most movies
director_counts[1, ]
```



Which director had the most movies by release year and remove the NA's?
```{r}
director_counts <- director_counts %>%
  filter(director != "NA")
director_counts[1, ]
```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix
```{r}
# Extract the top directors and their movie counts
top_directors <- head(director_counts, 10)  # Adjust the number as needed
top_director_names <- top_directors$director
top_director_counts <- top_directors$n

# Create a bar plot
barplot(top_director_counts, 
        names.arg = top_director_names,
        main = "Top Directors with Most Movies on Netflix",
        xlab = "Director",
        ylab = "Number of Movies",
        col = "skyblue",
        las = 2,  # Rotate x-axis labels vertically
        ylim = c(0, max(top_director_counts) + 2))  # Adjust ylim for better visualization
#difficlt to read
```

*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

I asked it to use ggplot

```{r}
# Load necessary library
library(ggplot2)

# Create a bar plot of the top directors with the most movies
ggplot(director_counts[1:10, ], aes(x = reorder(director, n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Director", y = "Number of Movies", title = "Top Directors with the Most Movies on Netflix") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()  # Flip the coordinates for horizontal bars

```

*How about for just 2020*



create a bar plot for the directors with the most movies just in 2020 using ggplot without creating a new object
```{r}
# Create a bar plot of the directors with the most movies in 2020
ggplot(netflix_movies, aes(x = director)) +
  geom_bar(data = subset(netflix_movies, format(date_added, "%Y") == "2020"), 
           stat = "count", fill = "skyblue") +
  labs(x = "Director", y = "Number of Movies", 
       title = "Directors with the Most Movies on Netflix in 2020") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
